### 2013-08-26 doc/dissertation/document/frontmatter/abstract.tex
moving estimator variance into an appendix

Predictive modeling and simulation is an important and developing
component of modern nuclear reactor analysis. New high performance
computing hardware and algorithms that leverage this hardware are
needed to help meet the goals and requirements of engineering
analysis. Current practices in neutron transport and fluid flow
leverage methods based on subspace iterations to achieve good
convergence properties and quality parallel scaling on large
machines. However, the next generation of machines is expected to
possess both high levels of concurrency and low memory availability
per core compared to today's machines with a potentially rate of
faults. Therefore, we seek new algorithms to solve these problems in
reactor analysis that have the potential to perform well on
forthcoming hardware with the goal of improving both iterative
performance and parallel scalability.

@@

Third, a new domain decomposed algorithm to parallelize MCSA aimed at
leveraging leadership-class computing facilities was developed by
utilizing parallel strategies from the radiation transport
community. The new algorithm utilizes the Multiple-Set
Overlapping-Domain strategy in an attempt to reduce parallel overhead
and add a natural element of replication to the algorithm. A large
number of parallel scaling studies using a neutron diffusion problem
were performed on the Titan Cray XK7 machine at Oak Ridge National
Laboratory with up to 65,356 cores used in the strong scaling studies
and 131,072 cores used in the weak scaling studies. It was found that
for the current implementation of MCSA, both weak and strong scaling
were superior to production implementations of Krylov methods. In
addition, the replication in the algorithm was able to boost parallel
efficiencies by up to 20\% at large core counts in the strong scaling
studies. By using the principle of superposition, replicating the
Monte Carlo sequence was also demonstrated to enhance time to
solution. Finally, it was discovered that MCSA exhibits the most
scalable behavior when it is formulated as a stochastic realization of
an additive Schwarz method.

Each of these research and development activities meet the goals of
this work by producing methods that in many cases were demonstrated to
converge in fewer iterations than conventional methods and exhibit
better parallel efficiencies at high levels of concurrency. All the
components of this work are required to enable future simulation of
nonlinear transport phenomena on the next generation of hardware. To
enable high performance simulations with FANM, we require the MCSA
techniques developed in the research on the neutron transport problem
and in turn, effectively leveraging MCSA on leadership-class hardware
requires the new parallel algorithm developed by this work. With the
further development of these methods based on this work, future
nuclear reactor simulations may potentially be enabled that have an
improved time to solution and more efficiently utilize the available
computational resources.

>>>>

Predictive modeling and simulation is an important and developing
component of modern nuclear reactor analysis. New high performance
computing hardware and algorithms that leverage this hardware are
needed to help meet the goals and requirements of engineering
analysis. Current practices in neutron transport and fluid flow
leverage methods based on subspace iterations to achieve good
convergence properties and quality parallel scaling on large
machines. However, the next generation of machines is expected to
possess both high levels of concurrency and low memory availability
per core compared to today's machines with a potentially high rate of
faults. Therefore, we seek new algorithms to solve these problems in
reactor analysis that have the potential to perform well on
forthcoming hardware with the goal of improving both iterative
performance and parallel scalability.

@@

Third, a new domain decomposed algorithm to parallelize MCSA aimed at
leveraging leadership-class computing facilities was developed by
utilizing parallel strategies from the radiation transport
community. The new algorithm utilizes the Multiple-Set
Overlapping-Domain strategy in an attempt to reduce parallel overhead
and add a natural element of replication to the algorithm. A large
number of parallel scaling studies using a neutron diffusion problem
were performed on the Titan Cray XK7 machine at Oak Ridge National
Laboratory with up to 65,356 cores used in the strong scaling studies
and 131,072 cores used in the weak scaling studies. It was found that
for the current implementation of MCSA, both weak and strong scaling
improved on that observed for production implementations of Krylov
methods. In addition, the replication in the algorithm was able to
boost parallel efficiencies by up to 20\% at large core counts in the
strong scaling studies and enhance time to solution in the weak
scaling stduies by using the principle of superposition. Finally, it
was discovered that MCSA exhibits the most scalable behavior when it
is formulated as a stochastic realization of an additive Schwarz
method.

Each of these research and development activities meet the goals of
this work by producing methods that in many cases were demonstrated to
converge in fewer iterations than conventional methods and exhibit
better parallel efficiencies at high levels of concurrency. All the
components of this work are required to enable future simulation of
nonlinear transport phenomena on the next generation of hardware. To
enable high performance simulations with FANM, we require the MCSA
techniques developed in the research on the neutron transport problem
and in turn, effectively leveraging MCSA on leadership-class hardware
requires the new parallel algorithm developed by this work. With the
further development of methods based on this work, future nuclear
reactor simulations may potentially be enabled that have an improved
time to solution and more efficiently utilize the available
computational resources.
